Supplement - Biserial correlation transformations

### 2019-07-15 code/Supplement-figures-tables.Rmd
Updated code file for supplement tables and figures

---
title: "Tables"
author: "Heterogeneity in direct replications in psychology and its association with effect size"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load packages, data and source functions}
if(!require(readr)){install.packages("readr")}
if(!require(metafor)){install.packages("metafor")}
if(!require(dplyr)){install.packages("dplyr")}
if(!require(purrr)){install.packages("purrr")}
if(!require(kableExtra)){install.packages("kableExtra")}

library(readr) #To load data
library(dplyr) #For data transformation
library(purrr) #For data iteration
library(metafor) #To run meta-analyses
library(kableExtra) #For creating good-looking pdf tables

@@

##Table 1
###Effect size $\rho_{xy}$ and its heterogeneity as a function of true effect size and measurement reliability.  
```{r effect-size-heterogeneity-table, cache = TRUE}
#library(kableExtra)
rooted_relabilities <- matrix(rep(c(.6, .7,.8), 3), ncol = 3)
rho <- c(0, .3, .5)
rel <- c("$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = .6", "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = .7", "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = .8")

rooted_relabilities <- sweep(rooted_relabilities, 2, rho, "*")
rooted_relabilities <- data.frame(rel, rooted_relabilities)

rooted_relabilities %>% knitr::kable(booktabs = T, col.names = c("", "$\\rho_{xy}$ = 0", "$\\rho_{xy}$ = .3", "$\\rho_{xy}$ = .5"), escape = FALSE, digits = 2) %>% 
  kable_styling(position = "left") %>% 
  footnote(threeparttable = TRUE, escape = FALSE,
           general = "Values in cells are observed effect sizes arising from the true effect size $\\\\rho_{xy}$ and measurement reliabilites $\\\\sqrt{R_{xx'}}\\\\times\\\\sqrt{R_{yy'}}$ when sample size is infinite. Increasing column-wise variance in observed effect sizes illustrates the association between effect size, reliability of measurement and heterogeneity  Code to reproduce table: osf.io/kf6pt/") %>% 
  row_spec(3, hline_after = T)

@@

##Table 2
###Pre-registered multi-lab replication projects
```{r data-summary-table}
#library(dplyr)
#library(kableExtra)

#Shortened version of APA reference
papers <- c('Klein et al. (2014)', 
'Ebersole et al. (2016)',
'Alogna et al. (2014)',
'Alogna et al. (2014)',
'Eerland et al. (2016)',
'Hagger et al. (2016)',
'Cheung et al. (2016)',
'Wagenmakers et al. (2016)',
'Bouwmeester et al. (2017)',
"O'Donnell et al. (2017)")

##Full APA reference below
# c('Klein, R. A., Ratliff, K. A., Vianello, M., Adams, R. B., Jr., Bahn?k, S., Bernstein, M. J., . . . Nosek, B. A. (2014). Investigating variation in replicability: A "many labs" replication project. Social Psychology, 45(3), 142-152.', 
# 'Ebersole, C. R., Atherton, O. E., Belanger, A. L., Skulborstad, H. M., Allen, J. M., Banks, J. B., ... & Brown, E. R. (2016). Many Labs 3: Evaluating participant pool quality across the academic semester via replication. Journal of Experimental Social Psychology, 67, 68-82.',
# 'Alogna, V. K., Attaya, M. K., Aucoin, P., Bahn?k, S., Birch, S., Birt, A. R., ... & Buswell, K. (2014). Registered replication report: Schooler and engstler-schooler (1990). Perspectives on Psychological Science, 9(5), 556-578.',
# 'Alogna, V. K., Attaya, M. K., Aucoin, P., Bahn?k, S., Birch, S., Birt, A. R., ... & Buswell, K. (2014). Registered replication report: Schooler and engstler-schooler (1990). Perspectives on Psychological Science, 9(5), 556-578.',
# 'Eerland, A., Sherrill, A. M., Magliano, J. P., Zwaan, R. A., Arnal, J. D., Aucoin, P., ... & Crocker, C. (2016). Registered replication report: Hart & Albarrac?n (2011). Perspectives on Psychological Science, 11(1), 158-171.',
# 'Hagger, M. S., Chatzisarantis, N. L., Alberts, H., Anggono, C. O., Batailler, C., Birt, A. R., ... & Calvillo, D. P. (2016). A multilab preregistered replication of the ego-depletion effect. Perspectives on Psychological Science, 11(4), 546-573.',
# 'Cheung, I., Campbell, L., LeBel, E. P., Ackerman, R. A., Aykutoglu, B., Bahn?k, S., ... & Carcedo, R. J. (2016). Registered Replication Report: Study 1 from Finkel, Rusbult, Kumashiro, & Hannon (2002). Perspectives on Psychological Science, 11(5), 750-764.',
# 'Wagenmakers, E. J., Beek, T., Dijkhoff, L., Gronau, Q. F., Acosta, A., Adams Jr, R. B., ... & Bulnes, L. C. (2016). Registered Replication Report: Strack, Martin, & Stepper (1988). Perspectives on Psychological Science, 11(6), 917-928.',
# 'Bouwmeester, S., Verkoeijen, P. P., Aczel, B., Barbosa, F., B?gue, L., Bra?as-Garza, P., ... & Evans, A. M. (2017). Registered Replication Report: Rand, Greene, and Nowak (2012). Perspectives on Psychological Science, 12(3), 527-542.',
# "O'Donnell, M., Nelson, L., McLatchie, N. M., & Lynott, D. J. (2017). Registered Replication Report: Dijksterhuis & van Knippenberg (1998): Multilab direct replication of: A variant of study 4 from Dijksterhuis, A., & van Knippenberg, A.(1998). The relation between perception and behavior, or how to win a game of trivial pursuit. Journal of personality and social psychology. Perspectives on Psychological Science."

description <- c("Two correlational effects: 'Gender math attitude' compared implicit attitudes (IAT) towards math between genders and 'IAT correlation math' correlated implicit attitudes with self-reported measures. The remainder were experiments with two independent groups. The groups were primed in some way (Anchoring 1-4; low vs. high category scales; norm of reciprocity; flag priming; currency priming), asked to imagine slightly different situations (Sunk costs; gain vs. loss framing; gambler's fallacy; imagined contact) or asked their agreement with statements presented differently (Allowed vs. forbidden; quote attribution). ", #ML1
            "Several effects were experiments with two independent groups. The groups were either primed in some way (Power and perspective; warmth perceptions; subjective distance interaction), saw slightly different statements (Elaboration likelihood interaction; credentials interaction) or experienced different situations (weight embodiment). Examined interactions were between treatment conditions and participant characteristics. One priming effect (metaphor) compared two treatment groups with a control. One effect was correlational: 'persistence and conscientiousness' was measured by an unsolvable anagram task and self-report respectively. The Stroop task is a within-person experiment with two conditions and the 'Availability' effect asks participants to judge whether some letters are more common in the first or third position.", #ML2
                 "Verbal overshadowing 1; Independent two-group experiment. Participants either described a robber after watching a video or listed countries/capitals and after a filler task attempted to identify the robber in a lineup.", #RRR1
            "Verbal overshadowing 2; Different from 1 only in that the filler task took place before the descriptive task instead of after.", #RRR2
                 "Grammar's effect on interpretation; Independent two-group vignette experiment with three outcome variables. Participants read about actions either described in imperfect or perfect tense and then rated protagonist's intentions (intentionality/intention attribution/detailed processing).", #RRR3
            "Ego depletion; Independent two-group experiment. Participants either assigned to a cognitively demanding task or a neutral, and performance was then measured in a subsequent cognitive task.", #RRR4
            "Commitment on neglect/exit; Independent two-group experiment with two outcome variables. Participants either primed to think about commitment to or independence from partner.", #RRR5
            "Facial feedback hypothesis; Independent two-group experiment. Participants either induced to 'smile' or 'pouth' by holding a pen in their mouth differently and simultaneously rated funniness of cartoons. ", #RRR6
            "Intuitive cooperation; Independent two-group experiment. Economic game with money contribution to a common pool either under time pressure or time delay.",
                 "Professor priming; Independent two-group experiment. Participants primed with either a 'professor' or 'hooligan' stimuli. Outcome was percentage correct trivia answers.") #RRR8

US_online <- dat %>% #Extract the (max when multiple effects) number of US/online labs per RP
  group_by(rp, effect) %>% 
  summarize(US = sum(country == "USA"),
            online = sum(in_lab == 0)) %>% 
  group_by(rp) %>% 
  summarize(US = max(US),
            online = max(online)) %>% 
  ungroup() %>% select(US, online)

sample_settings <- c("26/36 samples were primarily university students, 3 general population and 7 undescribed. 9/36 samples were online, including all the general population ones.", #https://osf.io/g3udn/, ML1
                     "20/21 samples were university students, 1 general population which was also the only online sample.", #Ml3
                     "31/32 samples were undergraduate students aged 18-25, 1 general population which was also the only online sample.", #RRR1
                     "22/23 samples were undergraduate students aged 18-25, 1 general population which was also the only online sample.", #RRR2
                     "11/12 samples were undergraduate students mostly aged 18-25, one of which was online. 1 sample was a broader online sample.", #RRR3
                     "All samples consisted of in-lab undergraduate students", #RRR4
                     "All samples consisted of in-lab undergraduate students aged 18-25", #RRR5
               "All but one sample explicitly consisted of students and all took place in-lab. The last sample was recruited at university grounds.", #RRR6
               'All samples consisted of in-lab undergraduate students aged 18-34.', #RRR7
                     "All samples consisted of in-lab undergraduate students aged 18-25") #RRR8

#1) what kind of samples

summary.table <- dat %>% 
  group_by(rp) %>% 
  summarize(Countries = n_distinct(country),
            k = n_distinct(Site), 
            Effects = n_distinct(effect),
            N = if(Effects == 1){sum(Ntotal)}
                           else{sum(Ntotal) / Effects}) %>% #average across effects
  mutate(Paper = papers,
         'Sample and Settings' = sample_settings,
         'Description of Effects' = description,
         'K (US)' = paste0(k, " (", US_online$US, ")")) %>% 
  select(RP = rp, Paper, Countries,'K (US)', everything(), -k) #change order of columns

summary.table %>% knitr::kable(booktabs = TRUE,
                               digits = 0,
                               linesep = "\\addlinespace",
                               longtable = TRUE) %>% 
  kable_styling(c("hold_position", "repeat_header"), font_size = 7) %>% 
  column_spec(2, width = "7em") %>% #specify width of paper column 
  column_spec(7, width = "16em") %>% 
  column_spec(8, width = "30em") %>% 
  landscape()
```
Note: "For studies with several effects the number of participants is the average across effects, rounded to the closest whole number. N = Participants used for primary analyses by original authors (i.e., after exclusions). RP = Replication Project, K (US) = no. primary studies (number of US studies), ML = Many Labs, RRR = Registered Replication Report. Code to reproduce table: osf.io/kf6pt/"

##Table 3
###Heterogeneity across primary effects and statistical power of ten multi-lab replication projects, ordered with respect to estimated between studies variance ($\tau^2$).
```{r prep-heterogeneity-table, cache = TRUE}
#library(metafor)
#library(dplyr)
#library(purrr)

>>>>

---
title: "Supplement - Biserial correlation transformations"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
Sys.setlocale("LC_TIME", "English" ) #to print date in YAML in English
```

```{r load packages, data and source functions}
if(!require(readr)){install.packages("readr")}
if(!require(metafor)){install.packages("metafor")}
if(!require(dplyr)){install.packages("dplyr")}
if(!require(purrr)){install.packages("purrr")}
if(!require(kableExtra)){install.packages("kablExxtra")}

library(readr) #To load data
library(dplyr) #For data transformation
library(purrr) #For data iteration
library(metafor) #To run meta-analyses
library(ggplot2) #plotting figures
library(cowplot) #combine plots
library(kableExtra) #create fancy pdf tables

@@

```{r a-plot-prep, cache = TRUE}
res <- dat %>% 
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map_dfr(est_heterogen_smd_raw, .id = "effect") #apply function, rbind results into dataframe ('dfr'), make sure purrr is up to date

het <- dat %>% 
  select(rp, Site, effect, effect_type) %>% 
  group_by(rp, effect) %>% 
  summarize(effect_type = unique(effect_type),
            k = n_distinct(Site)) %>% 
  left_join(res) %>% 
  mutate(effect_type = recode(effect_type, 
                              d = "SMD.",
                              'Raw mean difference' = "MD",
                              'Risk difference' = "RD")) %>% 
  ungroup()

a <- het %>% filter(rp == "ML1" & !effect %in% c('Allowed vs. forbidden', 'Gain vs. loss framing','Norm of reciprocity', 'Low vs. high category scales') & effect_type == "SMD.")
```

```{r, cache = FALSE}
a_plot <- ggplot(a, aes(x = s_I2, y = sqrt(tau2))) + geom_point() + ggtitle("SMD (untransformed)") +
  scale_x_continuous(name = expression(paste(italic(I)^2, "Index"))) +
  scale_y_continuous(name = "Tau")

@@

```{r trunc-biserial, cache = TRUE}
#Transform effect sizes to biserial correlations (pearson's correlations left as is)
#estimate tau2 for correlations
#'transform_MA' and 'summarizer' are sourced functions
res2 <- dat %>% 
  split(.$effect) %>%  #separate by effect, necessary step otherwise the function is applied overall
  map(transform_MA) %>% #
  map_dfr(summarizer, .id = "effect")

#Add rp and k to the data-frame 
het <- dat %>% 
  select(rp, Site, effect, effect_type) %>% 
  group_by(rp, effect) %>% 
  summarize(effect_type = unique(effect_type),
            k = n_distinct(Site)) %>% 
  left_join(res2) %>% 
  mutate(effect_type = recode(effect_type, 
                              d = "SMD.",
                              'Raw mean difference' = "MD",
                              'Risk difference' = "RD")) %>% 
  ungroup()

d <- het %>% filter(rp == "ML1" & !effect %in% c('Allowed vs. forbidden', 'Gain vs. loss framing','Norm of reciprocity', 'Low vs. high category scales') & effect_type == "SMD.") %>% 
  mutate(sigma2 = tau2/(I2/100) - tau2) #add 'typical within study variance' per project

```{r c-plot-1, cache = FALSE}
c_plot_1 <- ggplot(d, aes(x = I2, y = sqrt(tau2))) + geom_point() + ggtitle("Biserial r (transformed SMD)")

```{r untruncated-biserial, cache = TRUE}
res3 <- dat %>%
  filter(rp == "ML1" & !effect %in% c('Allowed vs. forbidden', 'Gain vs. loss framing','Norm of reciprocity', 'Low vs. high category scales') & outcomes1_2 == "mean _ SD") %>% 
  split(.$effect) %>% 
  map(~transform_SD(.$outcome_t1, .$outcome_c1, .$outcome_t2, .$outcome_c2, .$ntreatment, .$ncontrol)) %>% 
  map(~untruncated_d_to_r(.$d, .$n1, .$n2)) %>% 
  map_dfr(summarizer, .id = "effect")
```

```{r c-plot, cache = FALSE}
c_plot <- c_plot_1 + geom_point(data = res3, shape = 4, aes(x = I2, y = sqrt(tau2))) +
  scale_x_continuous(name = expression(paste(italic(I)^2, " Index (biserial)"))) +
  scale_y_continuous(name = "Tau (biserial)")

```{r tau-vs-tau}
b <- select(a, effect, tau2) %>% 
  left_join(select(d, effect, tau2), by = "effect") %>% #add truncated transformation tau2
  left_join(select(res3, effect, tau2), by = "effect") #add untruncated transformation tau2
```

```{r b-plot-1}
b_plot_1 <- ggplot(b, aes(x = sqrt(tau2.x), y = sqrt(tau2.y))) + 
  geom_point()

```{r b-plot}
b_plot <- b_plot_1 + geom_point(data = b, shape = 4, aes(x = sqrt(tau2.x), y = sqrt(tau2))) +
  ggtitle("SMD vs. biserial") +
  scale_x_continuous(name = "Tau (SMD)") +
  scale_y_continuous(name = "Tau (biserial)")
```

```{r, fig.width = 6, fig.asp = 1.3}
cowplot::plot_grid(a_plot, b_plot, c_plot, ncol = 1, labels = "AUTO")

# ggsave("../figures/Figure-S1.png", dpi = 600, width = 6, height = 7.8)
```

_Figure S.1_ Association between estimates of $\tau$ and $I^2$ after and before standardized mean differences (SMD) effects were transformed into biserial correlations for all effects reported as SMDs in ManyLabs 1. Panel A shows the association between estimates of $\tau$ and $I^2$ without transformation when using the meta-analytic specification of the original authors (see main text). Panel B shows the association between estimates of $\tau$ for the same effects before and after transformation to biserial correlations, either with truncation of correlations over 1 (circles) or without (crosses) when estimating variance. Panel C  shows the associations between estimates of $I^2$ on the SMD metric and estimates of $\tau$ after transformation, either with truncation or not.

#Table S1.
###Heterogeneity across primary effects and statistical power of ten multi-lab replication projects, ordered with respect to estimated heterogeneity ($I^2$).
```{r prep-heterogeneity-table, cache = TRUE}
#library(metafor)
#library(dplyr)
#library(purrr)
